Predicting and Preventing Employee Turnover through HR Analytics

### 2024-07-11 products/manuscript/Manuscript.qmd
Project part 3 - statistical analysis

save_data_location = here::here("data","processed-data","HRprocesseddata.rds") #using the here library for the location of new file
saveRDS(HRprocesseddata, file = save_data_location)
```

@@

### **Employee Tenure (in years)**

@@

**Tenure category**

@@

### **Gender**

@@

### **Monthly Income and Education**

@@

### **Age vs Attrition**

@@

## Additional Analysis

>>>>

save_data_location = here::here("data","processed-data","HRprocesseddata.rds") #using the here library for the location of new file
saveRDS(HRprocesseddata, file = save_data_location)

@@

### Employee Tenure (in years)

@@

### Tenure category

@@

### Gender

@@

### Monthly Income and Education

@@

### Age vs Attrition

@@

## Statistical Analysis Analysis

### Logistic Regression with Age

I conducted logistic regressions to explore predictors of Attrition among employees. Age emerged as a significant factor, with each additional year decreasing Attrition odds by approximately 0.052 (p < 0.001), underscoring its impact on turnover.

These findings emphasize the need for age-sensitive strategies, such as mentorship programs for knowledge transfer, flexible work arrangements, and tailored career development opportunities. Leveraging experience can foster commitment and reduce turnover among senior employees. However, limitations include the model's inability to capture all turnover influencers, like career aspirations or economic factors.

```{r}
#| label: tbl-logAge
#| tbl-cap: "Logistic Regression - Age"
#| echo: FALSE
logAge_table=readRDS("../../results/tables/logistic_age.rds")
knitr::kable(logAge_table)

### Logistic Regression with Income

Money plays a crucial role in life decisions, including work-related choices. I examined Monthly Income to understand its impact on Attrition among employees. The results show that higher Monthly Income is associated with lower odds of Attrition. Specifically, for each unit increase in Monthly Income, the odds of Attrition decrease by about 0.0001271 (p < 0.001), highlighting its significant influence on turnover.

To leverage these findings, organizations can consider implementing competitive salary structures, performance-based incentives, and financial wellness programs. These initiatives can enhance employee satisfaction and retention, particularly among those sensitive to income-related factors.
```{r}
#| label: tbl-logInc
#| tbl-cap: "Logistic Regression - Income"
#| echo: FALSE
logInc_table=readRDS("../../results/tables/logistic_income.rds")
knitr::kable(logInc_table)

### Logistic Regresion with Age, Income, and Job Satisfaction

I used a logistic regression model to study how Age, Monthly Income, and Job Satisfaction collectively affect Attrition among employees. This approach allowed me to explore their combined impact on turnover. The analysis revealed that older employees, those with higher incomes, and greater job satisfaction are significantly less likely to leave. Specifically, each additional year of Age reduces Attrition likelihood by about 0.0328 (p < 0.001), with similar significant impacts observed for Monthly Income and Job Satisfaction.
```{r}
#| label: tbl-glm
#| tbl-cap: "GLM - Age, Income, and Job Satisfaction"
#| echo: FALSE
glm_table=readRDS("../../results/tables/glm_model.rds")
knitr::kable(glm_table)
```

## Next Steps

We must investigate more predictors of attrition to identify root causes and develop essential strategies for retaining employees and improving the overall employee experience.
